*Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health

### 2023-05-05 3. Analysis Script/StatisticalAnalysis.qmd
Final Pushes of MADA

```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #For plotting/data vis.
library(broom) #For cleaning up output from lm()
library(dotwhisker) #Visualization for regression outputs
library(here) #Setting paths/for data loading/saving
library(tidymodels) #Great package for models and other resources used for fitting
library(tidyr) #Helps with data wrangling
library(gtsummary) #To create tables 
library(dplyr)
library(vip)
library(performance) #Model evaluation
library(jtools)
library(broom)
library(broom.mixed) #Converts fitted to tiddy dfs
```

@@

# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**

@@

# 3. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**

@@

tidy(lmall_PTGLM)
```

## a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**
Again, this is an alternative way to run a glm.

@@

# 4. **Data Fitting/Statistical Analysis: Generalized Linear Model [Feeder Trial]**

@@

check_model(lmall_FDGLM_all1$fit)
```

# **Generalized Linear Model: Feeder Data (Persistence)**

@@

FD_GLM_PREV
```

# **Time to event**

```{r}
library(finalfit)
library(forcats)
library(survival)

@@

PT_Data = PT_Data %>% 
  mutate(
      salmonella_crr= case_when(
        Salmonella_Positive ==  0~1
        Salmonella_Positive == 1~1
        TRUE ~ 1))

dependent_crr= "Surv(Date, status_crr)"
explanatory= c("Table", "Quadrant", "Sample_Type")

@@

PT_Data %>%
  filter(status==1) %>%
  summarize(median_surv=median(Date))
```

# **Cox regression: Picnic Table Data**

@@

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
  tbl_regression(exp=TRUE)
```

\#**Cox regression: Feeder Data**

```{r}
library(lubridate)

@@

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
  tbl_regression(exp=TRUE)
```

>>>>

```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #For plotting/data vis.
library(broom) #For cleaning up output from lm()
library(dotwhisker) #Visualization for regression outputs
library(here) #Setting paths/for data loading/saving
library(tidymodels) #Great package for models and other resources used for fitting
library(tidyr) #Helps with data wrangling
library(gtsummary) #To create tables 
library(dplyr)
library(vip)
library(performance) #Model evaluation
library(jtools)
library(broom)
library(broom.mixed) #Converts fitted to tiddy dfs
library(finalfit) #Needed for Time to Event Analysis 
library(forcats) #Needed for Time to Event Analysis 
library(survival) #Needed for Time to Event Analysis 
library(lubridate)
library(ggsurvfit)
library(gt)
library(webshot) #To save tables
library(knitr) #For tables
```

@@

# 2a. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**

@@

# 2b. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**

@@

tidy(lmall_PTGLM)
```
## 3a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**

Again, this is an alternative way to run a glm.

@@

# 3b. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Feeder Trial\]**

@@

check_model(lmall_FDGLM_all1$fit)
```
## 3b. **Generalized Linear Model: Feeder Data (Persistence)**

@@

FD_GLM_PREV
```
# 4a. **Time to Event Analysis: Picnic Table Trial**
```{r}

PT_Data_TTE<- #First let's format the dates 
  PT_Data %>% 
  mutate(
    Date= ymd(Date))

PT_Data_TTE$Date<- as.numeric(PT_Data_TTE$Date) #Let's make these numeric

PT_Data_TTE <- #I want to add a column displaying time between events (by day)
  PT_Data_TTE %>%
  mutate(
    os_day= as.duration(Date)/ddays(1))

PT_Data_TTE$os_day<- as.numeric(PT_Data_TTE$os_day) #Again let's make sure our new column is in the proper format

surv1<- Surv(PT_Data_TTE$Date,PT_Data_TTE$Salmonella_Positive)

#Let's visualize!
sfit1<- survfit(Surv(Date, Salmonella_Positive) ~  1, data=PT_Data_TTE)  %>% 
  ggsurvfit() + 
  labs(
    x = "Days",
    y = "Overall Salmonella Persistence") + #Let's add some labels
  add_confidence_interval() +
  add_risktable()

#I am not confident with this graphic so I will include this in my supplemental section
```

```{r}
glimpse(PT_Data)

PT_Data = PT_Data %>% 
 mutate(
  salmonella_crr= case_when(
   Salmonella_Positive ==  0~1,
    Salmonella_Positive == 1~1,
     TRUE ~ 1))

dependent_crr= "Surv(Date, salmonella_crr)"
explanatory= c("Table", "Quadrant", "Sample_Type")

@@

@@

PT_MedSurv<-PT_Data %>%
 filter(salmonella_crr==1) %>%
  summarize(median_surv=median(Date))

PT_MedSurv <- ((PT_MedSurv)/31556952000)*100

PT_Surv_Table<- gt(PT_MedSurv) %>%
   tab_header(
    title = "Median Survival of Salmonella",
    subtitle = ("01/02/2020 to 01/09/2020")) %>%
  cols_label(
    median_surv= "Median Salmonella Survival (days)")

PT_Surv_Table

PT_Surv_Table %>%
  saveRDS(here("4. Results", "Picnic_Table_Median_Survival.rds")) #To include in my manuscript, persistence seems to be roughly 5 days
```
# 4b. **Time to Event Analysis: Feeder Trial**
```{r}
FD_Data_TTE<- #First let's format the dates 
  FD_Data %>% 
  mutate(
    Date= ymd(Collection_Date))

FD_Data_TTE$Collection_Date<- as.numeric(FD_Data_TTE$Collection_Date) #Let's make these numeric

FD_Data_TTE <- #I want to add a column displaying time between events (by day)
  FD_Data_TTE %>%
  mutate(
    os_day= as.duration(Collection_Date)/ddays(1))

FD_Data_TTE$os_day<- as.numeric(FD_Data_TTE$os_day) #Again let's make sure our new column is in the proper format

surv1_FD<- Surv(FD_Data_TTE$Collection_Date,FD_Data_TTE$Absence_0_Presence_1)

#Let's visualize!
sfit1<- survfit(Surv(Collection_Date, Absence_0_Presence_1) ~  1, data=FD_Data_TTE)  %>% 
  ggsurvfit() + 
  labs(
    x = "Days",
    y = "Overall Salmonella Persistence") + #Let's add some labels
  add_confidence_interval() +
  add_risktable()

sfit1
```
```{r}
FD_Data = FD_Data %>% 
 mutate(
  salmonella_crr= case_when(
   Absence_0_Presence_1 ==  0~1,
    Absence_0_Presence_1 == 1~1,
     TRUE ~ 1))

dependent_crr= "Surv(Collection_Date, salmonella_crr)"
explanatory= c("Feeder_Type")

FD_Data %>% finalfit(dependent_crr, explanatory)

FD_Data %>% 
   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all)

survfit(Surv(Collection_Date, Absence_0_Presence_1)~1, data=FD_Data)

FD_MedSurv<-FD_Data %>%
 filter(salmonella_crr==1) %>%
  summarize(median_surv=median(Collection_Date))

FD_MedSurv <- ((FD_MedSurv)/604800)*100 #Persistence seems to be roughly 3 days for the FD

FD_Surv_Table<- gt(FD_MedSurv) %>%
   tab_header(
    title = "Median Survival of Salmonella",
    subtitle = ("02/22/2022 to 03/14/2022")) %>%
  cols_label(
    median_surv= "Median Salmonella Survival (days)")

FD_Surv_Table

FD_Surv_Table %>%
  saveRDS(here("4. Results", "Feeder_Median_Survival.rds")) #To include in my manuscript, seems to be roughly 3 days
```

# 5a. **Cox regression: Picnic Table Trial**

@@

CR_PT1<-coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
  tbl_regression(exp=TRUE)

CR_PT1

CR_PT1%>%
  saveRDS(here("4. Results", "PT_Cox.rds"))

```

# 5b. **Cox regression: Feeder Trial**

```{r}
FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)

@@

CR_FD1<- coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
  tbl_regression(exp=TRUE)

CR_FD1

CR_FD1%>%
  saveRDS(here("4. Results", "FD_Cox.rds"))
```
It looks like the

### 2023-05-05 5. Manuscript/Manuscript.qmd
Final Pushes of MADA

## **Description of data and data source**

@@

More packages were loaded that would assist in the visualization process
before begining this process. Before visualization began, the glimpse()
function was utilized to view the recently cleaned data. Given the
uncertainty of *Salmonella's* persistence on various surface materials
and the information these trials hoped to provide, the summary()
function was utilized to provide a quick glimpse of basic result
summaries for positive *Salmonella* samples and if the pile present was
the original pile or not. Exploration continued on these two variables
of interest by plotting each using a stacked bar graph via ggplot().

@@

Statistical analysis was conducted on both datasets, Picnic Table Data
and Feeder Data, utilizing RStudio "Cherry Blossom" (2023-03-09).
Analysis for both datasets were evaluated utilizing a linear regression
model.

@@

**Picnic Table Data Exploration and Visualization**
```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}

@@

**Bird Feeder Data Exploration and Visualization**
```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}

>>>>

### **Description of data and data source**

@@

More packages were loaded that would assist in the visualization process
before beginning this process. Before visualization began, the glimpse()
function was utilized to view the recently cleaned data. Given the
uncertainty of *Salmonella's* persistence on various surface materials
and the information these trials hoped to provide, the summary()
function was utilized to provide a quick glimpse of basic result
summaries for positive *Salmonella* samples and if the pile present was
the original pile or not. Exploration continued on these two variables
of interest by plotting each using a stacked bar graph via ggplot().

@@

Statistical analysis was conducted on the Picnic Table Data and Feeder
Data, utilizing RStudio "Cherry Blossom" (2023-03-09). Analysis for both
datasets were evaluated utilizing several models including Linear
Regression (LR), Generalized Linear Model (GLM), and Cox Regression
(CR). Given that two datasets were utilized for this analysis, a
holistic approach, considering various limitations of both trials, such
as sample size, was taken into account when selecting a model. Because
the majority of data is binary in nature (e.g., presence v. absences)
and few variables exist for comparison, a LR was the first model
selected to run on both datasets, followed by a GLM, and CR. Several
predictors were selected to include in this the analysis including
'Salmonella_Positive\~Date' , 'Salmonella_Positive\~Sample_Type' (e.g.,
pooled v. persistent), and 'Salmonella_Positive\~.' with all predictors
for the Picnic Table Trial and 'Absence_0\_Presence_1\~Feeder_number',
'Absence_0\_Presence_1\~Feeder_number+Feeder_Type', and
'Absence_0\_Presence_1\~.' and all predictors for the Feeder Trial.

Both Linear Regression and Generalized Linear Models performed as
expected when model performance was checked utilizing the check_models()
function. Given the flexibility and model performance of the GLM, it was
selected as the model for inclusion in our analysis between predictors
(e.g., 'Date' \[PT\], 'Sample_Type' \[PD\], 'Feeder_Type' \[FD\]) and
outcomes (e.g., 'Salmonella_Positive' \[PT\], 'Absence_0\_Presence_1'
\[FD\]). To

While LR and GLM provide insight into

@@

**Picnic Table Data Exploration and Visualization**

```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}

@@

**Bird Feeder Data Exploration and Visualization**

```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}
